Model based reinforcement learning at FERMI FEL

### 2020-11-16 main.tex
txs auto checkin

\section{Deep Reinforcement learning}
Assume states $s$ the set of all states $\states$ and actions $a$ all actions $\actions$ and rewards $r$ and the set of all rewards $\mathcal{R}$ an initial state distribution $d_0$.
Goal of reinforcement learning is to find a $\pi(s)\mapsto a$, which is the solution of:
\begin{align}
\max J(\pi)  =  \mathbb E_{\tau \thicksim p_{\pi}}\left[\sum_{t=0}^{H}\gamma^t r(s_t,a_t)\right],
\end{align}
where $p_\pi = d_0(s_0)\prod_{t_0}^H\pi(a_t,s_t)\mathcal T(s_{t+1}|s_t,a_t)$ is the distribution of all trajectories $\tau := (s_0, a_0, s_1, a_1,...s_H,a_H)$ drawn by $\pi$ with a horizon $H$.
In the modern field of RL one distinguishes if the policy $\pi(s)\approx \pi_\phi(s)$ or the state-action value function $Q(s,a)$ is approximated using a high capacity function approximator, as e.g. a deep neural network. In the first case one speaks about policy gradient methods in the latter about approximate dynamic programming, which we now discuss.
\subsection{Approximate dynamic programming}
The state-value function $Q(s,a)$ is expressed as $Q_\phi(s,a)$, where $\phi$ denotes the parameters of the function approximator. By satisfying the Bellmann-optimality operator the optimal $Q^*(s,a)$ can be found:
\begin{align}
    \min_\theta \left(\vec Q_\theta - \mathcal{B}^*\vec Q_\theta\right)^2.\label{eq:minimize_bellmann_optimality}
\end{align}
This operator has a unique fixed point but is non-linear, due to the $\max$ - operator:
\begin{align}
 \mathcal{B}^*\vec Q_\theta(s_t,a_t) = r(s_t,a_t)+\gamma\max_a\left( Q_\theta(s_{t+1},a_t) - Q_\theta(s_t,a_t)\right).
\end{align}
The form of this equation can cause overestimation and other complications, when using a function approximator. Several methods exist which try to mitigate the problems \cite{}.
One way to reduce the effect is to take a simple analytical form of the $Q$-function.
\subsection{NAF}
If a specific quadratic form of the $Q$ function is assumed:
\begin{align}
     Q_\theta(s,a) = -\frac{1}{2}(a-\mu_\theta(s))P_\theta(a-\mu_\theta(s))^T+V_\theta(s).
\end{align}
 The optimal policy is given by:
\begin{align}
    \pi^*(s)=(\delta(a)-\arg\!\max_a Q_\theta(s,a)).
\end{align}
One modification, which is used in these tests is a twin network (weights denoted by $\tilde\theta$), where only one is used to obtain the policy, while the other is used for the update rule to avoid over-estimation.
The maximum is given analytically as $\max_a Q(s,a) = V(s)$, hence from \cref{eq:minimize_bellmann_optimality}:
\begin{align}
	\min_\theta\left( (r(s_t,a_t)+\gamma \min_{\theta\prime,\tilde \theta} V_{\theta\prime}(s_{t+1}) - (1+\gamma) Q_\theta(s_t,a_t)\right)^2
\end{align}
$\theta\prime$ denotes the weights of a previous iteration.
%\subsection{New NAF network architecture}

\begin{algorithm}[ht]
\caption{On-policy policy gradient with Monte Carlo estimator \label{alg:pg}}
\begin{algorithmic}[1]
\State initialise $\theta_0$
\For{iteration $k \in [0, \dots, K]$}
%\State sample trajectories $\{\traj_i\}$ by running $\policy_{\theta_k}(\ba_t|\bs_t)$ \Comment{each $\traj_i$ consists of $\bs_{i,0},\ba_{i,0},\dots,\bs_{i,H},\ba_{i,H}$}
%\State compute $\return_{i,t} = \sum_{t'=t}^H \discount^{t'-t} \reward(\bs_{i,t},\ba_{i,t})$
%\State fit $b(\bs_t)$ to $\{\return{i,t}\}$ \Comment{use constant $b_t = \frac{1}{N}\sum_i \return{i,t}$, or fit $b(\bs_t)$ to $\{\return{i,t}\}$}
%\State compute $\hat{A}(\bs_{i,t},\ba_{i,t}) = \return_{i,t} - b(\bs_t)$
%\State estimate $\nabla_{\theta_k} J(\policy_{\theta_k}) \approx \sum_{i,t} \nabla_{\theta_k} \log \policy_{\theta_k}(\ba_{i,t} | \bs_{i,t}) \hat{A}(\bs_{i,t},\ba_{i,t})$
%\State update parameters: $\theta_{k+1} \leftarrow \theta_k + \alpha \nabla_{\theta_k} J(\policy_{\theta_k})$
\EndFor
\end{algorithmic}
\end{algorithm}

@@

\begin{figure}[!h]
	\centering
	\includegraphics*[width=0.5\textwidth]{Figures/MBRL_overview}
	\caption{A schematic overview of the AE-Dyna approach used in this paper.}
	\label{fig:MBRL_overview}
\end{figure}
\subsection{What was tested?}
\begin{itemize}
	\item The ANN - including the prior - number of models - noise level
	\item The number of data points and the policy
	\item The uncertainty - how is this included
	\item The episodic design
	\item ...
\end{itemize}
\subsection{Operational deployment}
\NB{Niky} % come abbiamo interagito con FERMI
\subsection{\NB{short FERMI introduction}}
The Free Electron laser Radiation for Multidisciplinary Investigation  better known as the FERMI, is the seeded free-electron laser facility next to the third-generation synchrotron facility ELETTRA at the Elettra Sinchrotrone Trieste laboratories.
A free-electron laser is a fourth-generation light source where the lasing medium consists of very-high-speed electron moving freely through a magnetic structure. The FERMI peculiarity is given by the usage of an external seeding source that provides several advantages, as the increased stability in pulse and photon energy, reduced size of the device, improved longitudinal coherence, more control on the pulse structure, and improved temporal synchronization with external timing systems.

>>>>

\section{Deep Reinforcement learning}
Assume states $s$ the set of all states $\states$ and actions $a$ all actions $\actions$ and rewards $r$ and the set of all rewards $\mathcal{R}$ an initial state distribution $d_0$.
Goal of reinforcement learning is to find a $\pi(s)\mapsto a$, which is the solution of:
\begin{align}
\max J(\pi)  =  \mathbb E_{\tau \thicksim p_{\pi}}\left[\sum_{t=0}^{H}\gamma^t r(s_t,a_t)\right],
\end{align}
where $p_\pi = d_0(s_0)\prod_{t_0}^H\pi(a_t,s_t)\mathcal T(s_{t+1}|s_t,a_t)$ is the distribution of all trajectories $\tau := (s_0, a_0, s_1, a_1,...s_H,a_H)$ drawn by $\pi$ with a horizon $H$.
In the modern field of RL one distinguishes if the policy $\pi(s)\approx \pi_\phi(s)$ or the state-action value function $Q(s,a)$ is approximated using a high capacity function approximator, as e.g. a deep neural network. In the first case one speaks about policy gradient methods in the latter about approximate dynamic programming, which we now discuss.
\subsection{Approximate dynamic programming}
The state-value function $Q(s,a)$ is expressed as $Q_\phi(s,a)$, where $\phi$ denotes the parameters of the function approximator. By satisfying the Bellmann-optimality operator the optimal $Q^*(s,a)$ can be found:
\begin{align}
    \min_\theta \left(\vec Q_\theta - \mathcal{B}^*\vec Q_\theta\right)^2.\label{eq:minimize_bellmann_optimality}
\end{align}
This operator has a unique fixed point but is non-linear, due to the $\max$ - operator:
\begin{align}
 \mathcal{B}^*\vec Q_\theta(s_t,a_t) = r(s_t,a_t)+\gamma\max_a\left( Q_\theta(s_{t+1},a_t) - Q_\theta(s_t,a_t)\right).
\end{align}
The form of this equation can cause overestimation and other complications, when using a function approximator. Several methods exist which try to mitigate the problems \cite{}.
One way to reduce the effect is to take a simple analytical form of the $Q$-function.
\subsection{NAF}
If a specific quadratic form of the $Q$ function is assumed:
\begin{align}
     Q_\theta(s,a) = -\frac{1}{2}(a-\mu_\theta(s))P_\theta(a-\mu_\theta(s))^T+V_\theta(s).
\end{align}
 The optimal policy is given by:
\begin{align}
    \pi^*(s)=(\delta(a)-\arg\!\max_a Q_\theta(s,a)).
\end{align}
One modification, which is used in these tests is a twin network (weights denoted by $\tilde\theta$), where only one is used to obtain the policy, while the other is used for the update rule to avoid over-estimation.
The maximum is given analytically as $\max_a Q(s,a) = V(s)$, hence from \cref{eq:minimize_bellmann_optimality}:
\begin{align}
	\min_\theta\left( (r(s_t,a_t)+\gamma \min_{\theta\prime,\tilde \theta} V_{\theta\prime}(s_{t+1}) - (1+\gamma) Q_\theta(s_t,a_t)\right)^2
\end{align}
$\theta\prime$ denotes the weights of a previous iteration. To stabilize the network training a small artificial noise is added to the actions in the update.
%\subsection{New NAF network architecture}

\begin{algorithm}[ht]
\caption{On-policy policy gradient with Monte Carlo estimator \label{alg:pg}}
\begin{algorithmic}[1]
\State initialise $\theta_0$
\For{iteration $k \in [0, \dots, K]$}
\State sample trajectories $\{\traj_i\}$ by running $\policy_{\theta_k}(\ba_t|\bs_t)$ \Comment{each $\traj_i$ consists of $\bs_{i,0},\ba_{i,0},\dots,\bs_{i,H},\ba_{i,H}$}
%\State compute $\return_{i,t} = \sum_{t'=t}^H \discount^{t'-t} \reward(\bs_{i,t},\ba_{i,t})$
%\State fit $b(\bs_t)$ to $\{\return{i,t}\}$ \Comment{use constant $b_t = \frac{1}{N}\sum_i \return{i,t}$, or fit $b(\bs_t)$ to $\{\return{i,t}\}$}
%\State compute $\hat{A}(\bs_{i,t},\ba_{i,t}) = \return_{i,t} - b(\bs_t)$
%\State estimate $\nabla_{\theta_k} J(\policy_{\theta_k}) \approx \sum_{i,t} \nabla_{\theta_k} \log \policy_{\theta_k}(\ba_{i,t} | \bs_{i,t}) \hat{A}(\bs_{i,t},\ba_{i,t})$
%\State update parameters: $\theta_{k+1} \leftarrow \theta_k + \alpha \nabla_{\theta_k} J(\policy_{\theta_k})$
%\EndFor
\end{algorithmic}
\end{algorithm}

@@

\begin{figure}[!h]
	\centering
	\includegraphics*[width=0.5\textwidth]{Figures/MBRL_overview}
	\caption{A schematic overview of the AE-Dyna approach used in this paper.}
	\label{fig:MBRL_overview}
\end{figure}
\subsection{What was tested?}
\begin{itemize}
	\item The ANN - including the prior - number of models - noise level.
	\item The number of data points and the policy.
	\item The uncertainty - how is this included.
	\item The episodic design.
	\item The MFRL agent as e.g. the TRPO and PPO or the SAC.
\end{itemize}
\subsection{Operational deployment}
\NB{Niky} % come abbiamo interagito con FERMI
\subsection{\NB{short FERMI introduction}}
The Free Electron laser Radiation for Multidisciplinary Investigation  better known as the FERMI, is the seeded free-electron laser facility next to the third-generation synchrotron facility ELETTRA at the Elettra Sinchrotrone Trieste laboratories.
A free-electron laser is a fourth-generation light source where the lasing medium consists of very-high-speed electron moving freely through a magnetic structure. The FERMI peculiarity is given by the usage of an external seeding source that provides several advantages, as the increased stability in pulse and photon energy, reduced size of the device, improved longitudinal coherence, more control on the pulse structure, and improved temporal synchronization with external timing systems.
